Modelling fine-scale dive behavior with hidden Markov models

### 2020-06-23 Report/Sections/model.tex
Update Report

One issue with the CarHMM is that it assumes Markovian dynamics conditioned on the hidden state, i.e. that any observation $Y^*_{t,t^*}$ depends only on the behavioral state $X^*_{t,t^*}$ and $Y^*_{t,t^*-1}$ (here we focus on the fine-scale process). However, there are many animal movement processes which violate this Markov property on very fine scales. For example, swimming behaviour of marine mammals can be periodic since the animal repeatedly flukes to propel itself forward. Work has been done in the past to model non-Markovian dynamics in the \textit{behavioural} process $X^*_t$ \cite{Langrock:2012}, but addressing non-Markovian dynamics within the observation process $Y^*_t$ is still a relatively unstudied area (MAYBE?). With improvements in tagging technology allowing for data collection at very high frequencies, data exhibiting noisy and non-Markovian fine scale behavior is likely to persist.

To address this issue, we recommend borrowing techniques from the signal processing literature to compress the data and summarize its essential elements. In particular, we suggest performing the discrete-time Short-time Fourier Transform (STFT) over each observed fine-scale process $Y^*_t$:
%
$$STFT\{Y^*_{t,t^*:t^*+w-1}\}(n) := \hat{Y}^{*(n)}_{t,t^*} = \sum_{n = 0}^{w-1} Y^*_{t,t^*+n}e^{-i \frac{2\pi k}{w} n} \qquad \forall n \in \{0, \ldots, w-1\}, \quad t^* \in \{1,w+1,2w+1, \ldots, w \left(\lfloor T^*_t / w \rfloor -1 \right) + 1\}.$$
%
The STFT slides a moving window of length $w$ across the time series $Y_t^*$ with a step size of $w$ and transforms the domain of each window from time to frequency. This allows the spectrum of $Y_t^*$ at time $t^*$ to be summarized by a $w$-dimensional vector of Fourier coefficients. While other step sizes can be used for the sliding window, we select $w$ to avoid serial dependence between windows.

If $Y^*_t \in \mathbb{R}^{T^*_t}$, then $\hat{Y}_t^* \in \mathbb{C}^{\lfloor T^*_t / w \rfloor \times w}$. Although this allows $Y^*_t$ to be represented in a way that eliminates obvious periodic behavior, the data set itself is still approximately as large as $Y^*_t$ itself. To reduce the size of $\hat{Y}^*_t$, we propose taking summary statistics of each window as follows:
%
$$Z_{t,t^*}^{*(1)} = \mathcal{R}\left(\hat{Y}^{*(0)}_{t,t^*}\right) \qquad Z_{t,t^*}^{*(2)} = \frac{1}{w}\sum_{n=1}^{\tilde{f}}|\hat{Y}^{*(n)}_{t,t^*}|^2$$
%
$Z_{t,t^*}^{*(1)}$ is equal to the average value of $Y_{t,t^*:t^*+w-1}^*$, and $Z_{t,t^*}^{*(2)}$ is equal to the squared 2-norm of the component of $Y_{t,t^*:t^*+w-1}^*$ that can be attributed to frequencies in the signal between $1$ and $\tilde{f}$ periods per window length. Both the window length $w$ and the max frequency $\tilde{f}$ are tuning parameters that should be tuned in a problem-specific way. $w$ should be long enough to capture the periodic behavior of the underlying process (at least as long as the length of a period), but short enough to avoid over-smoothing of the data and to maintain high resolution in the behavioral process $X^*$. $\tilde{f}$ should be selected such that the maximum frequency of $Y_t^*$ that makes biological sense is $\tilde{f}$ per window length. Note that these summary statistics are just one possible choice of summary statistics within each window, and future studies can adjust these definitions as needed. A visualization of transforming a one-dimensional sequence $Y^*$ to $Z^*$ can be seen in figure (\ref{fig:fourier_example}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=7.5in]{../Plots/fourier_transform.png}
	\caption{Visualization of transforming $Y^*_t$ into $Z^*_t$ using a sliding window and fourier transform.}
	\label{fig:fourier_example}
\end{figure}

Finally, note that it is possible to accommodate for unequal time steps within each window by using the \textbf{non-uniform discrete Fourier transform (NDFT)}. We do not describe the details of this method in this work, but the generalization is straightforward. Refer to Bagchi et al \cite{Bagchi:2001} for details.

@@

Hierarchical hidden Markov models can be used to jointly model simultaneous coarse-scale and fine-scale processes taking place simultaneously. However, as mentioned before, the fine-scale process $Y^*$ can often exhibit auto-correlation and intricate structure. Transforming $Y^*_t$ to $Z^*_t$ removes fine-scale periodic behavior, but $Z^*_t$ can still exhibit auto-correlation, especially in the window average, $Z_{t,t^*}^{*(1)}$. Therefore, we replace the fine-scale HMM within the Hierarchical HMM with a CarHMM according to figure (\ref{fig:CarHHMM}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=7.5in]{../Plots/CarHHMM.png}
	\caption{Graphical representation of a CarHHMM. The additional arrows representing auto-correlation between observations are shown in red for emphasis.}
	\label{fig:CarHHMM}
\end{figure}

The likelihood of this model is still easy to calculate using the forward algorithm:
%
$$\calL_{\text{CarHHMM}}(y,z^*;\theta,\theta^*,\Gamma,\Gamma^*,\delta,\delta^*) = \delta P(y_1,z_1^*;\theta,\theta^*,\Gamma^*,\delta^*) \prod_{t=2}^T \Gamma P(y_t,z_t^*;\theta,\theta^*,\Gamma^*,\delta^*) \mathbf{1}$$
%
where:
%
\begin{align*}
P(y_t,z_t^*;\theta,\theta^*,\Gamma^*,\delta^*)  = \text{diag}\left[p_{\theta}(y_t|X_t = x_1)\calL_{\text{CarHMM}}\left(z_t^*;\theta^{*(x_1)},\Gamma^{*(x_1)},\delta^{*(x_1)}\right), . . . , \right.\\
\left. p_{\theta}(y_t|x_t = x_N )\calL_{\text{CarHMM}}\left(z_t^*;\theta^{*(x_N)},\Gamma^{*(x_N)},\delta^{*(x_N)}\right) \right]
\end{align*}
%

>>>>

One issue with the CarHMM is that it assumes Markovian dynamics when conditioned on the hidden state, i.e. that any observation $Y^*_{t,t^*}$ depends only on the behavioral state $X^*_{t,t^*}$ and $Y^*_{t,t^*-1}$. However, there are many animal movement processes which violate this Markov property on very fine scales. For example, swimming behaviour of marine mammals often exhibits periodic behavior since the animal repeatedly flukes to propel itself forward. Work has been done in the past to model non-Markovian dynamics in the \textit{behavioural} process $X^*_t$ \cite{Langrock:2012}, but addressing non-Markovian dynamics within the observation process $Y^*_t$ is still a relatively unstudied area (MAYBE?). With improvements in tagging technology allowing for data to be collected at very high frequencies, noisy and non-Markovian fine scale behavior is likely to persist.

To address this issue, we recommend borrowing techniques from signal processing literature to compress the data and summarize its essential elements. In particular, we suggest performing the discrete-time short-time Fourier transform (STFT) over the observed fine-scale process $Y^*_t$:
%
\begin{align*}
    STFT\{Y^*_{t,(s^*w)+1:(s^*+1)w) }\}(k) := \hat{Y}^{*(k)}_{t,s^*} = \sum_{n = 1}^{w} Y^*_{t,(s^*w)+n}e^{-i \frac{2\pi k}{w} (n-1)} \\ k = 0, 1, \ldots, w-1; \quad s^* = 0,1, \ldots, \lfloor T^*_t / w \rfloor - 1
\end{align*}
%
Where $i$ in the equation above refers the $\sqrt{-1}$. Note here that $s^*$ indexes the \textit{windows} moving across $Y^*_t$ while $t^*$ indexes the \textit{observations} of $Y^*_t$.

The STFT slides a moving window of length $w$ across the time series $Y_t^*$ and transforms the domain of each window from time to frequency. This allows the spectrum of $Y_t^*$ at time $t^*$ to be summarized by a $w$-dimensional vector of Fourier coefficients. While other step sizes can be used for the sliding window, we select $w$ to avoid serial dependence between windows.

If $Y^*_t \in \mathbb{R}^{T^*_t}$, then $\hat{Y}_t^* \in \mathbb{C}^{\lfloor T^*_t / w \rfloor \times w}$. Although this allows $Y^*_t$ to be represented in a way that eliminates obvious periodic behavior, the data set itself is still approximately as large as $Y^*_t$ itself. To reduce the size of $\hat{Y}^*_t$, we take two summary statistics of each window:
%
$$Z_{t,s^*}^{*(1)} = \mathcal{R}\left(\hat{Y}^{*(0)}_{t,s^*}\right) \qquad Z_{t,s^*}^{*(2)} = \frac{1}{w}\sum_{k=1}^{\tilde{f}}|\hat{Y}^{*(k)}_{t,s^*}|^2$$
%
$Z_{t,s^*}^{*(1)}$ is equal to the average value of $Y_{t,s^*}^*$, and $Z_{t,s^*}^{*(2)}$ is equal to the squared 2-norm of the component of $Y_{t,s^*}^*$ that can be attributed to frequencies in the signal between $1$ and $\tilde{f}$ periods per window length. Both the window length $w$ and the max frequency $\tilde{f}$ are problem-specific tuning parameters. $w$ should be long enough to capture the periodic behavior of the underlying process (at least as long as the length of a period), but short enough to avoid over-smoothing of the data and to maintain high resolution in the behavioral process $X^*$. $\tilde{f}$ should be selected such that the maximum frequency of $Y_t^*$ that makes biological sense is $\tilde{f}$ per window length. Note that these summary statistics are just one possible choice to describe each window. Other choices of summary statistics include the maximum and argmax of $\hat Y^*_{s^*,t}$. Future studies can adjust these definitions as needed. A visualization of transforming a one-dimensional sequence $Y^*$ to $Z^*$ can be seen in figure (\ref{fig:fourier_example}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=5in]{../Plots/fourier_transform.png}
	\caption{Visualization of transforming $Y^*_t$ into $Z^*_t$ using a sliding window and fourier transform.}
	\label{fig:fourier_example}
\end{figure}

Note that it is possible to accommodate for unequal time steps within each window by using the \textbf{non-uniform discrete Fourier transform (NDFT)}. We do not describe the details of this method in this work, but the generalization is straightforward. Refer to Bagchi et al \cite{Bagchi:2001} for details.

Fourier analysis has previously been used in the field of animal movement to explain animal behavior \cite{Fehlmann:2017} and specifically fluking \cite{Shorter:2017} from accelerometer data, and Fourier analysis within an HMM has been used to describe daily behavioral cycles of marine mammals  \cite{Heerah:2017}, but we are currently unaware of any studies which have incorporated Fourier analysis of accelerometer data within the structure of an HMM.

@@

Hierarchical hidden Markov models can be used to jointly model simultaneous coarse-scale and fine-scale processes taking place simultaneously. However, as mentioned before, the fine-scale process $Y^*$ can often exhibit auto-correlation and intricate structure. Transforming $Y^*_t$ to $Z^*_t$ removes fine-scale periodic behavior, but $Z^*_t$ can still exhibit auto-correlation, especially in $Z_{t,s^*}^{*(1)}$. Therefore, we replace the fine-scale HMM within the Hierarchical HMM with a CarHMM, as shown in figure (\ref{fig:CarHHMM}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=5in]{../Plots/CarHHMM.png}
	\caption{Graphical representation of a CarHHMM. The additional arrows representing auto-correlation between observations are shown in red for emphasis.}
	\label{fig:CarHHMM}
\end{figure}

The likelihood of this model is still easy to calculate using the forward algorithm:
%
$$\calL_{\text{CarHHMM}}(y,z^*;\Theta,\Theta^*,\Gamma,\Gamma^*) = \delta P(y_1,z_1^*;\Theta,\Theta^*,\Gamma^*) \prod_{t=2}^T \Gamma P(y_t,z_t^*;\Theta,\Theta^*,\Gamma^*) \mathbf{1}_N$$
%
where:
%
\begin{align*}
P(y_t,z_t^*;\Theta,\Theta^*,\Gamma^*)  = \text{diag}\Big[&f^{(1)}(y_t)\calL_{\text{CarHMM}}\left(z_t^*;\Theta^{*(1)},\Gamma^{*(1)}\right), \ldots , \\
&f^{(N)}(y_t)\calL_{\text{CarHMM}}\left(z_t^*;\Theta^{*(N)},\Gamma^{*(N)}\right) \Big]
\end{align*}
%

### 2020-06-23 Report/Sections/simulation.tex
Revise Report
Update Report

To test the CarHHMM with STFT, 100 separate sequences of 100 marine-animal dives were simulated. The coarse-scale observations $Y$ were set as the duration of each dive, and the fine-scale observations $Y^*$ were set as one-dimensional acceleration readings simulated at 50 hertz. the specific procedure used to generate data was as follows:

@@

	\begin{align*}
		&\Gamma = \begin{pmatrix} 0.4 & 0.6 \\ 0.6 & 0.4 \end{pmatrix} \\
		% 
		&Y_t|X_t \sim Gamma \\
		&\bbE(Y_t|X_t = 1) = 15 s, \quad \quad \bbE(Y_t|X_t = 2) = 60 s \\
		%
		&\bbV(Y_t|X_t = 1) = 25 s^2, \quad \quad  \bbV(Y_t|X_t = 2) = 100 s^2
	\end{align*}

	\item Once the dive durations were calculated, for each $t \in \{1, \ldots, 100\}$, dive $t$ was broken into $\lfloor Y_t/2 \rfloor$ 2-second segments (the end of the dive sequence was discarded). Further, each 2-second segment was assigned a behaviour according to a fine-scale Markov chain $X^*_t$, where $X^*_{t,t^*} \in \{1,2\}$ and $t^* \in \{1,101,201,\ldots,100\lfloor Y_t/2 \rfloor - 99\}$. The parameters of the fine-scale behaviour Markov chain were set to be as follows:
	%
	$$\Gamma^{*(1)} = \begin{pmatrix} 0.25 & 0.75 \\ 0.75 & 025 \end{pmatrix} \qquad 	\Gamma^{*(2)} = \begin{pmatrix} 0.75 & 0.25 \\ 0.25 & 0.75 \end{pmatrix}$$
	%
	where $\Gamma^{*(1)}$ was used for dives where $X_t = 1$ and $\Gamma^{*(2)}$ was used for dives where $X_t = 2$

	\item For each 2-second segment, the Fourier modes $\hat{Y}^*_{t,t^*}$ were simulated using the following procedure. Note that the $n^{th}$ Fourier mode of $\hat{Y}^*_{t,t^*}$ is denoted as $\hat{Y}^{*(n)}_{t,t^*}$ :

	\begin{align*}
    	\hat{Y}^{*(0)}_{t,1} &\sim \mathcal{N} \left(\mu = 1, \sigma^2 = 0.0025 \right) & \\
    	%
    	\hat{Y}^{*(0)}_{t,t^*} &\sim \mathcal{N} \left(\mu = 0.9 Y^{*(0)}_{t,t^*-100} + 0.1, \sigma^2 = 0.0025 \right), & t^* \in \{101,201,\ldots, 100*\lfloor Y_t/2 \rfloor + 1\} \\
    	%
    	\hat{Y}^{*(n)}_{t,t^*} &= a_{t,t^*}^{(n)} i\sqrt{b^{(n)}_{t,t^*}}, & n \in \{1,\ldots,49\} \\
	\end{align*}
	\begin{align*}
	    a_{t,t^*}^{(n)} &\sim  \left\{\begin{array}{lr}
    	-1 & w.p. \enspace 1/2 \\
    	1  & w.p. \enspace 1/2
    	\end{array}\right. \\
    	%
    	(b^{(n)}_{t,t^*}|X^*_{t,t^*}  = 1) &\sim Gamma(1/n^2, 1) \\
    	%
    	(b^{(n)}_{t,t^*}|X^*_{t,t^*} = 2) &\sim \left\{\begin{array}{lr}
    	Gamma(1/n^2, 1) & \text{for } n \neq 2\\
    	Gamma(20,1) & \text{for } n = 2
    	\end{array}\right. 
	\end{align*}

    \begin{align*}
        \hat{Y}^{*(50)}_{t,t^*} &= 0 & \\
    	%
    	\hat{Y}^{*(n)}_{t,t^*}  &= -\hat{Y}^{*(100-n)}_{t,t^*}, & \qquad n \in \{51,\ldots,99\}
    \end{align*}

	\item Finally, $Y^*_{t,t^*:t^*+99}$ was set using the inverse discrete Fourier transform of $\hat{Y}^*_{t,t^*}$:
	$$Y^*_{t,t^*:t^*+99} = IDFT\left(\hat{Y}^*_{t,t^*}\right)$$

@@

There are several practical reasons behind this construction. First, $\hat{Y}^*_{t,t^*}$ is anti-symmetric about $\hat{Y}^{*(50)}_{t,t^*}$ so that its inverse Fourier transform is real-valued. $\hat{Y}^{*(n)}_{t,t^*}$ also decays like $1/n^2$ so that $Y^*_{t,t^*}$ remains continuous within a 2-second window. Note, however, that $Y^*_{t,t^*}$ is not continuous in $t^*$ \textit{between} windows, as can be seen in figure (\ref{fig:fourier_example}). However, note the jumps are not too severe since $\hat{Y}^{*(0)}_{t,t^*}$ and $\hat{Y}^{*(0)}_{t,t^*+100}$ are highly correlated. See figure (\ref{fig:sim_data}) for details.

@@

\begin{align*}
    \delta &= \begin{pmatrix} 0.5 & 0.5 \end{pmatrix}\\
	\Gamma &= \begin{pmatrix} 0.4 & 0.6 \\ 0.6 & 0.4 \end{pmatrix} \\
	% 
	Y_t|X_t = 1 &\sim Gamma\left(\mu = 15 s, \sigma = 5 s\right) \\
	Y_t|X_t = 2 &\sim Gamma\left(\mu = 60 s, \sigma = 10 s\right) \\
\end{align*}

@@

\begin{align*}
    \left(Z^{*(1)}_{t,t^*+100} | Z^{*(1)}_{t,t^*} \right) &\sim \mathcal{N} \left(\phi^* * Z^{*(1)}_{t,t^*} + \phi^* * \mu^*, \sigma^{*2} \right), & \mu^* = 1, \enspace \phi^* = 0.9, \enspace \sigma^{*2} = 0.025 \\
    %
    \left(Z^{*(2)}_{t,t^*} | X^*_{t,t^*} = 1\right) &\sim Gamma\left(\alpha^{*(1)}, \beta^* \right), & \alpha^{*(1)} = \sum_{n=1}^{\tilde{f}} \frac{1}{n^2}, \enspace \beta^* = 1 \\
    %
    \left(Z^{*(2)}_{t,t^*} | X^*_{t,t^*} = 2\right) &\sim Gamma\left(\alpha^{*(2)}, \beta^* \right), & \alpha^{*(2)} = 20 + \sum_{n=2}^{\tilde{f}} \frac{1}{n^2}, \enspace \beta^* = 1 \\
\end{align*}
%
The emission distribution of the 2-second average acceleration $Z_{t,t^*}^{*(1)}$ in particular can be reparameterized as an OU process:
%
$$dZ^{*(1)}_{t,t^*} = \beta^*(\gamma^* - Z^{*(1)}_{t,t^*})dt^* + \omega^* dW_{t,t^*}$$

$$\gamma^* = \mu^* = 1, \qquad \beta^* = -\frac{\log(\phi^*)}{\Delta t^*} = 0.0229, \qquad \omega^{*2} = \frac{2\sigma^{*2}\log(\phi)}{\Delta t (\phi^{*2}-1)} = 0.006$$

Note that the two fine-scale states $X^*_{t,t^*} = 1$ and $X^*_{t,t^*} = 2$ are identical except for the fact that the distribution of $Z^{*(2)}_{t,t^*}$ is shifted up in its shape parameter in fine-scale state 2. This corresponds to more periodic behavior in acceleration, but \textit{not} a higher mean acceleration.

@@

Four different models were fit to the 100 simulated data sets:

\begin{enumerate}
    \item A CarHMM with no hierarchical component using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (CarHMM).
    \item A Hierarchical hidden Markov model with no auto-correlation using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (HHMM).
    \item A CarHHMM using $Z^{(1)}$, but not $Z^{(2)}$, as fine-scale observations (CarHHMM1).
    \item A CarHHMM using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (CarHHMM2).
\end{enumerate}
%
For each model, the following were recorded:
%
\begin{enumerate}
    \item an empirical distribution of the MLE estimates of $\Gamma$ and $\theta$ using the 100 generated data sets generated from true parameters $\Gamma_0$ and $\theta_0$.
    \item a confusion matrix showing the state decoding capabilities of each model across all data sets.
    \item The true coverage of 95\% confidence intervals estimated using the observed information matrix (to see how over- or under- confident our parameter estimates are).
\end{enumerate}

>>>>

To test the CarHHMM with STFT, 500 separate sequences of 100 marine-mammal dives were simulated. The coarse-scale observations $Y$ were set as the duration of each dive, and the fine-scale observations $Y^*$ were set as one-dimensional acceleration readings simulated at 50 hertz. the specific procedure used to generate data was as follows:

@@

	\begin{align*}
		&\Gamma = \begin{pmatrix} 0.4 & 0.6 \\ 0.6 & 0.4 \end{pmatrix} & \\
		% 
		&Y_t|X_t \sim \rm{Gamma} & \\
		&\bbE(Y_t|X_t = 1) = 15 s &\bbE(Y_t|X_t = 2) = 60 s \\
		%
		&\bbV(Y_t|X_t = 1) = 25 s^2 &\bbV(Y_t|X_t = 2) = 100 s^2
	\end{align*}

	\item Once the dive durations were calculated, for each $t \in \{1, \ldots, 100\}$, dive $t$ was broken into $\lfloor Y_t/2 \rfloor$ 2-second segments (the end of the dive sequence was discarded). Further, each 2-second segment was assigned a behaviour according to a fine-scale Markov chain $X^*_t$, where $X^*_{t,s^*} \in \{1,2\}$ and $s^* \in \{0,1,\ldots,\lfloor Y_t/2 \rfloor\}$. The parameters of the fine-scale behaviour Markov chain were set to be as follows:
	%
	$$\Gamma^{*(1)} = \begin{pmatrix} 0.25 & 0.75 \\ 0.75 & 025 \end{pmatrix} \qquad 	\Gamma^{*(2)} = \begin{pmatrix} 0.75 & 0.25 \\ 0.25 & 0.75 \end{pmatrix}$$
	%
	where $\Gamma^{*(1)}$ was used for dives where $X_t = 1$ and $\Gamma^{*(2)}$ was used for dives where $X_t = 2$

	\item For each 2-second segment, the Fourier modes $\hat{Y}^*_{t,s^*}$ were simulated using the following procedure. Note that the $k^{th}$ Fourier mode of $\hat{Y}^*_{t,s^*}$ is denoted as $\hat{Y}^{*(k)}_{t,s^*}$ :

	\begin{align*}
    	\hat{Y}^{*(0)}_{t,0} &\sim \mathcal{N} \left(\mu = 0, \sigma^2 = 0.0025 \right) & \\
    	%
    	\hat{Y}^{*(0)}_{t,s^*} &\sim \mathcal{N} \left(\mu = 0.9Y^{*(0)}_{t,s^*-1}, \sigma^2 = 0.0025 \right), & s^* = 1,2,\ldots, \lfloor Y_t/2 \rfloor \\
    	%
    	\hat{Y}^{*(k)}_{t,t^*} &= a_{t,s^*}^{(k)} i\sqrt{b^{(k)}_{t,s^*}}, & k = 1,\ldots,49 \\
	\end{align*}
	\begin{align*}
	    a_{t,s^*}^{(k)} &\sim  \left\{\begin{array}{lr}
    	-1 & \quad w.p. \enspace 1/2 \\
    	1  & \quad w.p. \enspace 1/2
    	\end{array}\right. \\
    	%
    	(b^{(k)}_{t,s^*}|X^*_{t,s^*}  = 1) &\sim {\rm{Gamma}}(1/k^2, 1) \\
    	%
    	(b^{(k)}_{t,t^*}|X^*_{t,t^*} = 2) &\sim \left\{\begin{array}{lr}
    	{\rm{Gamma}}(1/k^2, 1), \quad & k \notin \{1,2\} \\
    	{\rm{Gamma}}(100,1), & k = 1 \\
    	{\rm{Gamma}}(50,1), & k = 2
    	\end{array}\right. 
	\end{align*}

    \begin{align*}
        \hat{Y}^{*(50)}_{t,s^*} &= 0 & \\
    	%
    	\hat{Y}^{*(k)}_{t,s^*}  &= -\hat{Y}^{*(100-n)}_{t,s^*}, & \qquad k = 51,\ldots,99
    \end{align*}

	\item Finally, $Y^*_{t,(100s^*+1):(100s^*+100)}$ was set using the inverse discrete Fourier transform of $\hat{Y}^*_{t,s^*}$:
	$$Y^*_{t,(100s^*+1):(100s^*+100)} = IDFT\left(\hat{Y}^*_{t,s^*}\right)$$

@@

There are several practical reasons behind this construction. First, $\hat{Y}^*_{t,t^*}$ is anti-symmetric about $\hat{Y}^{*(50)}_{t,s^*}$ so that its inverse Fourier transform is real-valued. $\hat{Y}^{*(k)}_{t,s^*}$ also decays like $1/k^2$ so that $Y^*_{t,t^*}$ remains continuous within a 2-second window. Note, however, that $Y^*_{t,s^*}$ is not continuous \textit{between} windows, as can be seen in figure (\ref{fig:fourier_example}). However, the jumps are not too severe since $\hat{Y}^{*(0)}_{t,s^*}$ and $\hat{Y}^{*(0)}_{t,s^*+1}$ are highly correlated. See figure (\ref{fig:sim_data}) for details.

@@

\begin{align*}
    \delta &= \begin{pmatrix} 0.5 & 0.5 \end{pmatrix}\\
	\Gamma &= \begin{pmatrix} 0.4 & 0.6 \\ 0.6 & 0.4 \end{pmatrix} \\
	% 
	Y_t|X_t = 1 &\sim {\rm{Gamma}}\left(\mu = 15 s, \sigma = 5 s\right) \\
	Y_t|X_t = 2 &\sim {\rm{Gamma}}\left(\mu = 60 s, \sigma = 10 s\right) \\
\end{align*}

@@

\begin{align*}
    \left(Z^{*(1)}_{t,s^*} | Z^{*(1)}_{t,s^*-1} \right) &\sim \mathcal{N} \left(\phi^*Z^{*(1)}_{t,s^*-1} + \phi^*\mu^*, \sigma^{*2} \right), & \mu^* = 0, \enspace \phi^* = 0.9, \enspace \sigma^{*2} = 0.025 \\
    %
    \left(Z^{*(2)}_{t,s^*} | X^*_{t,s^*} = 1\right) &\sim {\rm{Gamma}}\left(\alpha^{*(1)}, \beta^* \right), & \alpha^{*(1)} = \sum_{k=1}^{\tilde{f}} \frac{1}{k^2}, \enspace \beta^* = 1 \\
    %
    \left(Z^{*(2)}_{t,t^*} | X^*_{t,s^*} = 2\right) &\sim {\rm{Gamma}}\left(\alpha^{*(2)}, \beta^* \right), & \alpha^{*(2)} = 20 + \sum_{k=2}^{\tilde{f}} \frac{1}{k^2}, \enspace \beta^* = 1 \\
\end{align*}
%
The emission distribution of the 2-second average acceleration $Z_{t,s^*}^{*(1)}$ can be reparameterized as an OU process according to Theorem 1:
%
$$dZ^{*(1)}_{t,t^*} = \beta^*(\gamma^* - Z^{*(1)}_{t,t^*})dt^* + \omega^* dW_{t,t^*}$$

$$\gamma^* = \mu^* = 0, \qquad \beta^* = -\frac{\log(\phi^*)}{\Delta t^*} = 0.0229, \qquad \omega^{*2} = \frac{2\sigma^{*2}\log(\phi)}{\Delta t (\phi^{*2}-1)} = 0.006$$

@@

Four different models were fit to each simulated data set:

\begin{enumerate}
    \item A CarHMM with no hierarchical component using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (CarHMM).
    \item A Hierarchical hidden Markov model with no auto-correlation using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (HHMM).
    \item A CarHHMM using $Z^{(1)}$, but not $Z^{(2)}$, as fine-scale observations (CarHHMM1).
    \item A CarHHMM using both $Z^{(1)}$ and $Z^{(2)}$ as fine-scale observations (CarHHMM2).
\end{enumerate}
%
For each model, the following were recorded:
%
\begin{enumerate}
    \item An empirical distribution of the MLE estimates of $\Gamma$ and $\Theta$ using the 100 generated data sets generated from true parameters $\Gamma_0$ and $\Theta_0$.
    \item The estimated spread of these distributions using the observed information matrix.
    \item A confusion matrix showing the state decoding capabilities of each model across all data sets.
\end{enumerate}

The empirical distribution of
